---
title: "vignetteDownscaler"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
packages <- c( 'viridis','rnaturalearthhires',
              'sf','INLA','rgeoboundaries','tidyverse')
package.check <- lapply(packages, FUN = function(x){
  if(!require(x ,character.only = T)){install.packages(x)}
  if(!(x %in% ( .packages()  ) ) ){library(x ,character.only = T)}
  })
theme_set(theme_minimal())
options(ggplot2.continuous.colour="viridis")
options(ggplot2.continuous.fill = "viridis")
```

```{r}
source("fnCheckInputsDown.R")
source("fnCreateMesh.R")
source("fnPredictDown.R")
```

# Specifying the arguments of `fnPredictDown()`

`fnPredictDown()` aims to predict spatial data when areal data and point data within the observed areal data available. The prediction is made by the following model:
$$
Y(s_i) = \beta_0(s_i) + \beta_1(s_i)X(s_i) + \epsilon_i
$$
where $Y(s_i)$ is the point data and $X(s_i)$ is the observed areal data that the point $Y(s_i)$ locates. 

Thus, To use `fnPredictDown()`, the observed point (`depoint`) and areal data (`dearea`) are compulsory. They are `sf` objects. 

Then, we need to specify the locations or areas where we want to get predictions. 
Depending on where we want to predict,we would need to specify `sf` objects`dppoint` for a set of points and `dparea` for a set of areas.

`dppoint` can be an `sf` object containing a few points or a dense set of regular points within the study region representing a continuous surface
(`dpcontsurface`).

We also need to specify an `sf` object the boundary of the region of study, `boundaryregion`.

Note that all objects need to be `sf` objects and have the same projection.
Therefore, if we have `raster` or `sp` data, we need to transform them to `sf` to be able to use the package.

## Specifying the observed data


Similar to the vignette of `fnPredictMelding()`, we will show how to predict PM 2.5 in UK. The projection method is 4326 or WSG84 (latitude and longitude).
 
We obtain the boundary of the region of study with the `rgeoboundaries` package and transform it to the chosen projection. Then we construct `sf` objects from the observed point data and areal data. 
Please read the previous vignette (Link), if you do not know how to construct `sf` objects and specify the estimated data.  

The column names of `depoint` data have to be `value`,`geometry`, where `value` stands for the estimated value for point data and the other column includes spatial information.

The column names of `dearea` data have to be `value`,`geometry`, where `value` stands for the estimated value for areal data and the other column includes spatial information.
```{r}
# CRS projection
crsproj <- 4326
library(rgeoboundaries)
boundaryregion <- geoboundaries("United Kingdom")
boundaryregion <- st_transform(boundaryregion , crsproj)[,6]
# observed areal data
dearea <- st_read("data/dearea.shp")
# observed point data
depoint <- read.csv("data/pointdata.csv")
depoint <- depoint[, c(2, 3, 4)]
colnames(depoint) <- c('value', 'y', 'x')
depoint <- depoint %>% st_as_sf(coords = c("x", "y"), dim = "XY") %>%
  st_set_crs(crsproj) %>% st_cast("MULTIPOINT")
colnames(dearea) <- c('value', 'geometry')
```

## Specifying the point data and areal data for prediction

Areal data for prediction `dparea` is an `sf` object with columns `value` and `geometry`.
The estimated point position and area will be the predicted point and area. 

```{r}
dppoint <- depoint
dparea <- dearea
```

If we wish to predict in a continuous surface, we need to provide a dense grid of points within the region of study. The method can be found (Link). 

## Downscaler Method with `fnPredictDown()`

We use downscaler method only when both areal data and point data available. In this case, the mesh is the same, gnereated by combination of the location of the estimated point data and the boundary of the region. The predicted objectives can be point data, areal data, or a continuous surface. 

The following example shows how to predict both point data and areal data. Here the locations of the point data and the partition of the areal data is the same as the estimated objects. We are able to only predict point data or areal data by specifying the not required obj to be NULL.

```{r}
# Create mesh
mesh <- fnCreateMesh(depoint, boundaryregion)
plot(mesh)
points(as.matrix(st_coordinates(depoint)[ , c(1, 2)]), col = 2)
respre <- fnPredictDown(depoint = depoint, dearea = dearea, dppoint = dppoint, dparea = dparea, boundaryregion = boundaryregion, mesh = mesh)
head(respre)
  
ggplot(data = boundaryregion) + geom_sf() +
  geom_sf(data = respre[[1]], aes(geometry = geometry, color = pred_mean))+
  labs(title = "Average PM 2.5  Level 2016, UK", fill = "PM 2.5")
ggplot(data = boundaryregion) + geom_sf() +
  geom_sf(data = respre[[2]], aes(geometry = geometry, fill = pred_mean))+
  labs(title = "Average PM 2.5  Level 2016, UK", fill = "PM 2.5")
```

#### Predicting a continuous surface

Similar to Bayesian Melding method, we can use Downscaler method to predict a continuous surface and use the plot function `plot_sp()`. Here, we use the same mesh that we create in last section. 

```{r}
respre_down <- fnPredictDown(depoint = depoint, dearea = dearea, dppoint = dpcontsurface, dparea = NULL, boundaryregion = boundaryregion, mesh = mesh)

results = respre_down[[1]]

plot_sp(results)
```
